---
title: "Session 2: Cancer in the world around us"
shorttitle: "Rates and maps of cancer"
author: "Vincent Carey"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Session 2: Cancer in the world around us}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    number_sections: yes
    theme: united
    toc: yes
---

```{r setup, results="hide", echo=FALSE}
suppressPackageStartupMessages({
library(BiocYES)
library(sf)
library(tibble)
library(DT)
library(dplyr)
library(curatedTCGAData)
library(TCGAutils)
library(ontoProc)
})
```

# Some illustrative maps of cancer incidence



# Cancer map concepts

## Drawing maps with R

```{r lkd1}
library(BiocYES)
mass_map()
```

Exercise:  Look at the code for `mass_map` and change it to 
produce a map of Rhode Island.

## Tabulating cancer statistics with geographic metadata

### Some basic statistics by "metropolitan statistical area (MSA)"

```{r lkd2}
library(tibble)
library(dplyr)
data(woncan_meta)
cat(woncan_meta, sep="\n")
data(woncan)
littab = woncan |> select(MSA, `Cancer Sites`, Age.Adjusted.Rate) |> as.data.frame()
datatable(littab, options=list(lengthMenu=c(5,10,15,20)))
```

Exercise: Use the search input widget to limit focus to prostate cancer.  How many
MSAs do not report the age adjusted rate for this cancer?  Sort the rates in
decreasing order.  What is the MSA with second-highest rate of prostate cancer?  What
is the rate there?

### Representing locations

```{r lkd3}
library(sf)
data(us_county_geo)
us_county_geo[1,] |> select(state, county, geometry)
```

### Exercise

Join the location data to the `woncan` statistics.  Plot the
rates for prostate cancer at the geographic locations where
they were recorded.

As an example, look at the code for `cancer_map_usa`, which
produces:
```{r domaca}
cancer_map_usa()
```

