---
title: "Overview of Cancer Data Science concepts for YES for CURE"
shorttitle: "Intro to cancer data science for YES"
author: "Vincent Carey"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Overview of Cancer Data Science concepts for YES for CURE}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    number_sections: yes
    theme: united
    toc: yes
---

```{r setup, results="hide", echo=FALSE}
suppressPackageStartupMessages({
library(BiocYES)
library(sf)
library(tibble)
library(DT)
library(dplyr)
library(curatedTCGAData)
library(TCGAutils)
library(ontoProc)
})
```

# Goals of Cancer Data Science modules of YES for CURE

- Explain how data on cancer cases and deaths are collected and summarized in the United States
    - What are the roles of the Centers for Disease Control, National Cancer Institute, the states, territories and localities in gathering and reporting data?
    - Where can you go to see information about cancer trends of interest to you?
- Explain how cancer patients can learn about – and tell about – treatments and clinical trials relevant to their condition
- Explain how research data on the biology of cancer can be surveyed and explored from a "data science" perspective

"Data" plays a big part in thinking about our future engagement with cancer – as people, scientists, social leaders.  "Data science" is the study of how scientific discipline, including measurement, experimentation, and argument, guides us to use data effectively.  Data science is a young and diverse enterprise!


# A personal cancer data project

In 60 years of life I have been cancer free, but many of my relatives and friends have been
impacted directly:

```{r lkd}
library(BiocYES)
library(DT)
data(vjc_cancer_net)
datatable(vjc_cancer_net)
```

How can we interpret this?
